---
title: "ISL_Ch3_Linear Regression_exercise"
date: "2026-07-19"
output:
  md_document: default
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---
# Applied 

```{r, include=FALSE}
library(MASS)
library(ISLR2)
library(car)
```

## Exercise 8

```{r}
head(Auto)
```

Y: mpg, 每加侖可行駛英里數（Miles per Gallon）

X: horsepower, 馬力

$\hat{Y} = 39.935861 - 0.157845X$

```{r}
lm.fit = lm(mpg ~ horsepower, data = Auto)
summary(lm.fit)
```

### Confidence Interval and Prediction Interval

```{r}
confint(lm.fit)

predict(lm.fit, data.frame(horsepower = c(98)),
        interval = "prediction")

```


### Plot

```{r}
plot(y = Auto$mpg, x = Auto$horsepower)
abline(lm.fit, lwd = 3, col = "red")
```


### Diagnostic plots

```{r}
par(mfrow = c(2, 2))
plot(lm.fit)
```

Residuals vs Fitted 有點U型彎曲，故以下加入 $horsepower^2$

This suggests that the linearity assumption may be violated. Therefore, a quadratic term is added.

```{r}
lm.fit.2 = lm(mpg ~ horsepower + I(horsepower^2), data = Auto)
par(mfrow = c(2, 2))
plot(lm.fit.2)
```


## Exercise 9

### Scatterplot Matrix

```{r}
pairs(Auto[1:8])
```

### Correlation matrix

```{r}
names(Auto)
cor(Auto[1:8])
```

### Multiple Regression

```{r}
lm.fit.full = lm(mpg ~. -name ,data = Auto)
summary(lm.fit.full)
```

### Diagnostic plots

```{r}
par(mfrow = c(2, 2))
plot(lm.fit.full)
```

```{r}
which.max(cooks.distance(lm.fit.full))
max(cooks.distance(lm.fit.full))
```

### Interaction terms 

```{r}
names(Auto)
lm.fit.interaction <- lm(
  mpg ~ (cylinders + displacement + horsepower + weight +
           acceleration + year + origin)^2,
  data = Auto
)

#只看兩兩交互作用的效果 ^2 (pairwise interaction)
summary(lm.fit.interaction)
```

### Transformations of the variables

Try a few diﬀerent transformations of the variables, such as log(X), √X, X^2. Comment on your findings.

x: horsepower

y: mpg

sqrt : $Y = \beta _0 + \beta _1\sqrt{X} + \epsilon$

log : $Y = \beta _0 + \beta _1\ln{X} + \epsilon$

squared: $Y = \beta _0 + \beta _1 X^2  + \epsilon$

poly: $Y = \beta _0 + \beta _1 X_1 + \beta_2 X_2^2  + \epsilon$


```{r}
lm.fit.sqrt = lm(mpg ~ sqrt(horsepower), data = Auto)
lm.fit.log = lm(mpg ~ log(horsepower), data = Auto)
lm.fit.squared = lm(mpg ~ I(horsepower^2), data = Auto)
lm.fit.poly.squared = lm(mpg ~ horsepower + I(horsepower^2), data = Auto)
```


```{r}
par(mfrow = c(1,3))
plot(
  sqrt(Auto$horsepower), Auto$mpg,
  xlab = "sqrt(horsepower)",
  ylab = "mpg",
  main = "Square-root transformation"
)
abline(lm.fit.sqrt, lwd = 3, col = "red")

plot(
  log(Auto$horsepower), Auto$mpg,
  xlab = "log(horsepower)",
  ylab = "mpg",
  main = "Log transformation"
)
abline(lm.fit.log, lwd = 3, col = "blue")

plot(
  Auto$horsepower^2, Auto$mpg,
  xlab = "horsepower^2",
  ylab = "mpg",
  main = "Squared transformation"
)
abline(lm.fit.squared, lwd = 3, col = "green")
```


```{r}
par(mfrow = c(2, 2))

plot(
  fitted(lm.fit.sqrt),
  rstudent(lm.fit.sqrt),
  xlab = "Fitted values",
  ylab = "Studentized residuals",
  main = "sqrt(horsepower)"
)
abline(h = 0, lty = 2)

plot(
  fitted(lm.fit.log),
  rstudent(lm.fit.log),
  xlab = "Fitted values",
  ylab = "Studentized residuals",
  main = "log(horsepower)"
)
abline(h = 0, lty = 2)

plot(
  fitted(lm.fit.squared),
  rstudent(lm.fit.squared),
  xlab = "Fitted values",
  ylab = "Studentized residuals",
  main = "horsepower^2"
)
abline(h = 0, lty = 2)

plot(
  fitted(lm.fit.poly.squared),
  rstudent(lm.fit.poly.squared),
  xlab = "Fitted values",
  ylab = "Studentized residuals",
  main = "poly(horsepower, 2)"
)
abline(h = 0, lty = 2)

```

```{r}
summary(lm.fit.sqrt)$adj.r.squared
summary(lm.fit.log)$adj.r.squared
summary(lm.fit.squared)$adj.r.squared

sigma(lm.fit.sqrt)
sigma(lm.fit.log)
sigma(lm.fit.squared)
```


## Exercise 13

$X \sim N(0,1)$

$\epsilon \sim N(0,0.25)$

$Y = -1 + 0.5X +\epsilon$

```{r}
set.seed(1)
x = rnorm(n = 100, mean = 0, sd = 1)
eps = rnorm(n = 100, mean = 0, sd =sqrt(0.25))
y = -1 + 0.5 * x + eps
```

$\beta _ 0 = -1, \ \beta _1 =0.5$

$\hat{\beta _ 0} = -0.97578 , \ \hat{\beta _1} =0.55311$


```{r}
lm.fit.13 = lm(y ~ x)
summary(lm.fit.13)
```

### Plot

```{r}
plot(x = x, y = y)
abline(lm.fit.13,
       lwd = 3,
       col = "red")

abline(a = -1,
       b = 0.5,
       col = "blue",
       lwd = 3)

legend("topleft",
       legend = c("True line",
                  "Estimated line"),
       col = c("blue", "red"),
       lwd = 3)
```

### Polynomial regression model

Reduced model: $Y = \beta_0 + \beta_1X+\epsilon$

Full model: $Y = \beta_0 + \beta_1X+\beta_2X^2 +\epsilon$


Note: 

I(x^2) fits the raw polynomial,

poly(x, 2) uses orthogonal polynomials.


```{r}
lm.fit.13.poly = lm(y ~ x + I(x^2))
#note: I(x^2) 與 poly(x, 2)不同
summary(lm.fit.13.poly)
```

Partial F test: $H_0 :\beta_2=0$

Since the p-value = .722 > .05, we do not reject $H_0$.

There is no evidence that adding the quadratic term significiantly improves the model.


```{r}
anova(lm.fit.13, lm.fit.13.poly)
```

### Increasing/Decreasing the variance of epsilon

Repeat (a)–(f) after modifying the data generation process in such a way that there is more noise in the data. The model (3.39) should remain the same. You can do this by increasing the variance of the normal distribution used to generate the error term ϵ in (b). Describe your results.

1. $\epsilon '\sim N(0,\ 0.05)$

$Y = -1 + 0.5X +\epsilon'$


2. $\epsilon ''\sim N(0,\ 0.45)$

$Y = -1 + 0.5X +\epsilon''$


$\beta _ 0 = -1, \ \beta _1 =0.5$

V(ε) = 0.25 : $\hat{\beta _ 0} = -0.97578 , \ \hat{\beta _1} =0.55311$

V(ε) = 0.05 : $\hat{\beta _ 0}' = -0.99388 , \ \hat{\beta _1}' =0.50473$

V(ε) = 0.45 : $\hat{\beta _ 0}'' = -0.96132 , \ \hat{\beta _1}'' =0.46264$

```{r}
#smaller error variance
set.seed(1)
eps.05 = rnorm(n = 100, mean = 0, sd =sqrt(0.05))
y.05 = -1 + 0.5 * x + eps.05

lm.fit.le = lm(y.05 ~ x)
summary(lm.fit.le)


#greater error variance
eps.45 = rnorm(n = 100, mean = 0, sd =sqrt(0.45))
y.45 = -1 + 0.5 * x + eps.45
lm.fit.ge = lm(y.45 ~ x)
summary(lm.fit.ge)
```


### Plot (Comparison under Different Error Variances)

```{r}
par(mfrow = c(1, 3))
plot(x = x, y = y, main = "V(ε) = 0.25")
abline(lm.fit.13,
       lwd = 3,
       col = "red")

abline(a = -1,
       b = 0.5,
       col = "blue",
       lwd = 3)

legend("topleft",
       legend = c("True line",
                  "Estimated line"),
       col = c("blue", "red"),
       lwd = 3)

plot(x = x, y = y.05, main = "V(ε) = 0.05")
abline(lm.fit.le,
       lwd = 3,
       col = "red")

abline(a = -1,
       b = 0.5,
       col = "blue",
       lwd = 3)

legend("topleft",
       legend = c("True line",
                  "Estimated line"),
       col = c("blue", "red"),
       lwd = 3)

plot(x = x, y = y.45, main = "V(ε) = 0.45")
abline(lm.fit.ge,
       lwd = 3,
       col = "red")

abline(a = -1,
       b = 0.5,
       col = "blue",
       lwd = 3)

legend("topleft",
       legend = c("True line",
                  "Estimated line"),
       col = c("blue", "red"),
       lwd = 3)

```

As the error variance increases, 

1. the data become more scattered.

2. the estimated coefficients become less precise.

3. the fitted regression line deviates more from the population regression line.

### Confidence Interval

```{r}
confint(lm.fit.13)

confint(lm.fit.le)

confint(lm.fit.ge)
```

## Exercise 15

y : crim 每人平均犯罪率（城鎮犯罪程度）

x :

1. crim	犯罪率	

2. zn	大地住宅比例	Numeric	面積超過 25,000 平方英尺住宅用地比例

3. indus	非零售商業用地比例	Numeric	商業／工業發展程度

4. nox	一氧化氮濃度, 空氣污染程度

5. rm	平均房間數, 每戶平均房間數

6. age	老房屋比例, 1940 年以前建成住宅比例

7. dis	就業中心距離, 到五個 Boston 就業中心的加權距離

8. rad	高速公路便利性, 徑向高速公路可達性指數

9. tax	房屋稅率, 每 $10,000 房屋價值需繳的稅

10. ptratio	師生比, 城鎮學生／教師比例

11. lstat	低社經地位人口比例

12. medv	房價中位數

```{r}
vars = names(Boston)[names(Boston) != "crim"]
result = data.frame(
  Variable = vars,
  beta1 = NA,
  R2 = NA,
  pvalue = NA
)
#seq_along(x) = 1:length(x)
for(i in seq_along(vars)){
  formula_text = paste("crim ~", vars[i])   #建lm()公式的string
  formula_obj = as.formula(formula_text)   #把string變成lm()的公式
  fit = lm(formula_obj, data = Boston)
  result$R2[i] = summary(fit)$r.squared
  result$pvalue[i] = summary(fit)$coefficients[2, 4]
  result$beta1[i] = summary(fit)$coefficients[2, 1]
}


result[order(result$R2,decreasing = T), ] 
#order() return 排序之後的index

 #crim, zn, indus, chas, nox, rm, age, dis, rad, tax, ptratio, lstat, medv
```

All predictors except 'chas' show statistically significant associations with 'crim' at the 5% significance level.

The strongest linear association is observed for 'rad'($R^2=0.391$), followed by 'tax' and 'lstat'.

```{r}
par(mfrow = c(2,2))
plot(x = Boston$rad, y = Boston$crim)
plot(x = Boston$medv, y = Boston$crim)
plot(x = Boston$lstat, y = Boston$crim)
plot(x = Boston$rm, y = Boston$crim)
```


### Multiple Regression and Partial F Test

```{r}
full.model = lm(crim ~ .,data = Boston)
summary(full.model)
```

Partial F test

$H_0 : \beta_j=0$

In the multiple regression model, after controlling for all other predictors, rad, medv, dis, and zn are statistically significant at the 5% level. Therefore, we reject $H_0 : \beta_j=0$ for these four predictors.

Compared with the results in part (a), some variables, such as tax, lstat, nox have relatively high $R^2$ in the simple regression model. However, these variables are no longer statistically significant in the multiple regression model.


```{r}
result2 = data.frame(
  ExcludeVariable = vars,
  PartialF = NA,
  pvalue = NA
)

for(i in seq_along(vars)){
  formula_text = paste("crim ~. -", vars[i])  #建lm()公式的string, 除了vars[i]以外的變數都納入
  formula_obj = as.formula(formula_text)   #把string變成lm()的公式
  fit = lm(formula_obj, data = Boston)

  #names(anova(fit,full.model))
  result2$PartialF[i] = round(anova(fit,full.model)[2,"F"],4)
  result2$pvalue[i] = round(anova(fit,full.model)[2, "Pr(>F)" ],4)
}
result2[order(result2$pvalue , decreasing = F),]
```

### Plot(simple regression coefficient vs multiple regression coefficient)

The largest change occurred for 'nox', whose coefficient changed from 31.25 in the simple regression model to −9.96 in the multiple regression model. This sign reversal suggests strong multicollinearity with other predictors. Several other predictors (e.g., 'rm', 'ptratio', 'indus', 'zn', and 'age') also exhibited sign changes, although the magnitude of these changes was relatively smaller.

```{r}
mul.reg.coef = unname(summary(full.model)$coefficients[-1,"Estimate"])
coef.result = cbind(result[,1:2],
                    mul.reg.coef)

coef.diff = abs(coef.result[,2] - coef.result[,3]) #看回歸係數之差異絕對值
coef.sign.change = sign(coef.result[,2]) != sign(coef.result[,3])#看兩回歸係數是否有變號

coef.result = cbind(coef.result,
                    coef.diff,
                    coef.sign.change)
names(coef.result) = c("Variable", "S.R.coef", "M.R.coef", "Coef.diff", "Coef.sign.change")

coef.result[order(coef.result$Coef.diff, decreasing = T),]
```

```{r}
plot(x = coef.result$S.R.coef, 
     y = coef.result$M.R.coef)

text(
    coef.result$S.R.coef,
    coef.result$M.R.coef,
    labels = vars,
    pos = 4,
    cex = 0.8
)

abline(0,1,col="red",lty=2)
```

### non-linear association

full model: $Y = \beta_0+\beta_1X+\beta_2X^2 + \beta_3X^3+\epsilon$

reduced model: $Y = \beta_0+\beta_1X+\epsilon$

Partial F test

$H_0 : \beta_2 = \beta_3 = 0$

For every predictor, the partial F test rejected the null hypothesis $H_0 : \beta_2 = \beta_3 = 0$. This indicates that adding quadratic $\beta_2$ and cubic $\beta_3$  terms significantly improves the fit compared with a simple linear regression model, suggesting that evidence of nonlinear relationships between each predictor and 'crim'.


```{r}
vars = names(Boston)[
  names(Boston) != "crim" &
  names(Boston) != "chas"
]
vars
result3 = data.frame(
  Variable = vars,
  beta1 = NA,
  beta2 = NA,
  beta3 = NA,
  Fvalue = NA,
  pvalue = NA
)
#seq_along(x) = 1:length(x)
for(i in seq_along(vars)){
  formula_text = paste("crim ~", vars[i],
                       "+I(",vars[i],"^2)",
                       "+I(",vars[i],"^3)")   #建lm()公式的string
  formula_obj = as.formula(formula_text)   #把string變成lm()的公式
  fit = lm(formula_obj, data = Boston)
  
  formula_text.sim = paste("crim ~", vars[i])   
  formula_obj.sim = as.formula(formula_text.sim)   
  fit.sim = lm(formula_obj.sim, data = Boston)
  
  result3$beta1[i] = summary(fit)$coefficients[2, "Estimate"]
  result3$beta2[i] = summary(fit)$coefficients[3, "Estimate"]
  result3$beta3[i] = summary(fit)$coefficients[4, "Estimate"]
  result3$Fvalue[i] = anova(fit.sim, fit)[2,"F"]
  result3$pvalue[i] = round(anova(fit.sim, fit)[2,"Pr(>F)"],4)
}
result3[order(result3$pvalue, decreasing = F), ]
```




